---
title: "MapBiomasClassification"
output: 
  html_document: 
    keep_md: yes
---

This script analyses all years for a single classification

##Load Libraries
```{r warning=F, message=F}

rm(list=ls())
library(tidyverse)
library(raster)
library(readxl)
library(scales)
library(knitr)

```

##Functions
```{r}

#raster to xyz  (with help from https://stackoverflow.com/a/19847419)
#sepcify input raster, whether nodata cells should be output, whether a unique cell ID should be added
#return is a matrix. note format is row (Y) then col (X)
extractXYZ <- function(raster, nodata = FALSE, addCellID = TRUE){
  
  vals <- raster::extract(raster, 1:ncell(raster))   #specify raster otherwise dplyr used
  xys <- rowColFromCell(raster,1:ncell(raster))
  combine <- cbind(xys,vals)
  
  if(addCellID){
    combine <- cbind(1:length(combine[,1]), combine)
  }
  
  if(!nodata){
    combine <- combine[!rowSums(!is.finite(combine)),]  #from https://stackoverflow.com/a/15773560
  }
  
  return(combine)
}


getLCs <- function(data)
{
  #calculates proportion of each LC in the muni (ignoring NAs, help from https://stackoverflow.com/a/44290753)
  data %>%
    group_by(muniID) %>%
    dplyr::summarise(LC1 = round(sum(map == 1, na.rm = T) / sum(!is.na(map)), 3),
                     LC2 = round(sum(map == 2, na.rm = T) / sum(!is.na(map)), 3),
                     LC3 = round(sum(map == 3, na.rm = T) / sum(!is.na(map)), 3),
                     LC4 = round(sum(map == 4, na.rm = T) / sum(!is.na(map)), 3),
                     LC5 = round(sum(map == 5, na.rm = T) / sum(!is.na(map)), 3),
                     NonNAs = sum(!is.na(map)),
                     NAs = sum(is.na(map))
    ) -> LCs

  return(LCs)
}


```



```{r}

unzip(zipfile="MapBiomas_23_ASCII_unclassified_allYears.zip")  # unzip all file 


classification <- read_excel("MapBiomas_CRAFTY_classifications.xlsx", sheet = "PastureA", range="B2:C21", col_names=F)  

  
mb_data <- read_csv("LandCover Data - MapBiomas - Collection 2.3 - 2018.01.04 Municipios.csv")  

unzip(zipfile="sim10_BRmunis_latlon_5km_2018-04-27.zip",files="sim10_BRmunis_latlon_5km_2018-04-27.asc",exdir="ASCII")  # unzip file 
munis.r <- raster("ASCII/sim10_BRmunis_latlon_5km_2018-04-27.asc")  #do this with zip file

#extract cell values to table format
munis.t <- extractXYZ(munis.r, addCellID = F)
munis.t <- as.data.frame(munis.t)
munis.t <- plyr::rename(munis.t, c("vals" = "muniID"))

```


Loop over years creating CData table for all years

```{r}
yrls <- seq(2000,2015,1)

for(i in seq_along(yrls)){

  print(yrls[i])
  
  map <- raster(paste0("ASCII/brazillc_",yrls[i],"_5km_int.txt"))

  map <- reclassify(map, rcl=as.matrix(classification))
  plot(map, col=c("green", "pink", "yellow", "grey", "brown"))
  
  
  #extract cell values to table format
  map.t <- extractXYZ(map, addCellID = F)
  map.t <- as.data.frame(map.t)
  map.t <- plyr::rename(map.t, c("vals" = "map"))

  #so need to join 
  map_munis <- left_join(as.data.frame(munis.t), as.data.frame(map.t), by = c("row" = "row", "col" = "col"))

  #now summarise by muniID
  lcs_map_munis <- getLCs(map_munis)

  #convert cell counts to areas (km2) and add state id
  map_areas_munis <- lcs_map_munis %>%
    mutate(LC1area = round(LC1 * NonNAs) * 25) %>%
    mutate(LC2area = round(LC2 * NonNAs) * 25) %>%
    mutate(LC3area = round(LC3 * NonNAs) * 25) %>%
    mutate(LC4area = round(LC4 * NonNAs) * 25) %>%
    mutate(LC5area = round(LC5 * NonNAs) * 25) %>%
    mutate(state = substr(muniID, 1, 2))

  #drop original cell-count columns (work with area km2 from now on)
  map_areas_munis <- map_areas_munis %>% dplyr::select(-LC1, -LC2, -LC3, -LC4, -LC5, -NonNAs, -NAs)

  #summarise muni areas to state level
  map_areas <- map_areas_munis %>%
    group_by(state) %>%
    dplyr::summarise_at(vars(LC1area:LC5area),sum, na.rm=T) %>%  #use _at so state is not summarised
    mutate_if(is.character, as.integer)

  #gather to long format for union below
  map_areas <- map_areas %>%
    gather(key = LCa, value = area, -state)
  
  #recode LCs for union below
  map_areas <- map_areas %>%
    mutate(LC = if_else(LCa == "LC1area", 1, 
      if_else(LCa == "LC2area", 2,
      if_else(LCa == "LC3area", 3,
      if_else(LCa == "LC4area", 4,
      if_else(LCa == "LC5area", 5, 0)
      )))))

  #add source variable for plotting below (re-order to match map table for union below)
  map_areas <- map_areas %>%
    dplyr::select(-LCa) %>%
    mutate(source = "Map") %>%
    dplyr::select(state, LC, source, area)

  ###Summarise MapBiomas data to the states we are simulating

  #filter to get only the states we want
  mb_areas <- mb_data %>%
    filter(Estados == "TOCANTINS" | 
        Estados == "BAHIA" |
        Estados == "MINAS GERAIS" |
        Estados == "SÃO PAULO" |
        Estados == "PARANÁ" |
        Estados == "SANTA CATARINA" |
        Estados == "RIO GRANDE DO SUL" |
        Estados == "MATO GROSSO DO SUL" |
        Estados == "MATO GROSSO" |
        Estados == "GOIÁS")
  
  #add state column containing state ids
  mb_areas <- mb_areas %>%
    mutate(state = if_else(Estados == "TOCANTINS", 17, 
      if_else(Estados == "BAHIA", 29,
      if_else(Estados == "MINAS GERAIS", 31,
      if_else(Estados == "SÃO PAULO", 35,
      if_else(Estados == "PARANÁ", 41,
      if_else(Estados == "SANTA CATARINA", 42,
      if_else(Estados == "RIO GRANDE DO SUL", 43, 
      if_else(Estados == "MATO GROSSO DO SUL", 50, 
      if_else(Estados == "MATO GROSSO", 51,
      if_else(Estados == "GOIÁS", 52, 0
      ))))))))))
    )
  
  #!need to think about what tables to include year in...
  #!change needed here to access the year
  
  #select only columns we want    
  mb_areas <- mb_areas %>%
    dplyr::select(state, paste0(yrls[i]), `Classe Nivel 3`) %>% 
    dplyr::rename(area = paste0(yrls[i]))
  
  #because there is no consistency between land cover labels
  legenda = c("Forest Formations"=1, "Natural Forest Formations"=2, "Dense Forest"=3, "Savanna Formations"=4, "Mangroves"=5,"Forest Plantations"=9, "Non-Forest Natural Formations"=10, "Non Forest Wetlands"=11, "Grasslands"=12, "Other Non Forest Natural Formations"=13, "Farming"=14, "Pasture"=15, "Agriculture"=18, "Agriculture or Pasture"=21, "Non-Vegetated Areas"=22, "Dunes and Beaches"=23, "Urban Infrastructure"=24,"Other Non-Vegetated Area"=25,"Water Bodies"=26, "Non-Observed"=27)
  
  #recode to values (which match the map)
  mb_areas <- mb_areas %>%
    mutate(LC = recode(`Classe Nivel 3`, !!!legenda))
  
  #use the classification values from above to relassify land covers
  mb_areas$LC <- plyr::mapvalues(mb_areas$LC, from=as.numeric(classification$X__1), to=as.numeric(classification$X__2))
  
  #calculate total LC area by state
  mb_areas <- mb_areas %>%
    group_by(state, LC) %>%
    dplyr::summarise_at(vars(area),sum, na.rm=T) 
  
  #round to integer
  mb_areas <- mb_areas %>%
    mutate(area_km2 = round(area,0))
  
  #add source variable for plotting below (re-order to match map table for union below)
  mb_areas <- mb_areas %>%
    dplyr::select(-area) %>%
    mutate(source = "MB") %>%
    rename(area = area_km2) %>%
    dplyr::select(state, LC, source, area)
    
  
  #create CData table just for this year
  CData_yr <- union_all(map_areas, mb_areas)

  #relabel states to characters
  CData_yr <- CData_yr %>%
    mutate(state = if_else(state == 17, "TO", 
      if_else(state == 29, "BA",
      if_else(state == 31, "MG",
      if_else(state == 35, "SP",
      if_else(state == 41, "PR",
      if_else(state == 42, "SC",
      if_else(state == 43, "RS", 
      if_else(state == 50, "MS",
      if_else(state == 51, "MT",
      if_else(state == 52, "GO", "NA"
      ))))))))))
    )
  
  #relabel LCs to characters
  CData_yr <- CData_yr %>%
    mutate(LC = if_else(LC == 1, "Nature", 
      if_else(LC == 2, "OtherAgri",
      if_else(LC == 3, "Agri",
      if_else(LC == 4, "Other",
      if_else(LC == 5, "Pasture", "NA"
      )))))
    )
  
  
  #add year column
  CData_yr <- CData_yr %>%
    mutate(year = yrls[i])
  
  
  #union CData for years here.
  #if first iteration of loop create the tibble
  if(i == 1){
      CData <- CData_yr
  }
    
  #else join data to tibble (by creating another tibble, then join (ensure rows are not lost)
  else {
      CData <- union_all(CData, CData_yr)
  }
  
}  
```


Plot areas by state for a given year

```{r}  

yr <- 2000

#filter (remove) missing LCs
CData_yr <- CData %>%
  dplyr::filter(year == yr) %>%
  dplyr::filter(area > 0)

ggplot(CData_yr, aes(x=source, y=area, fill=LC)) + 
  geom_bar(stat="identity", colour="white", position = "fill") +
  scale_y_continuous(labels = percent_format()) +
  facet_grid(.~state) +
  xlab("Data Source")+ylab("Percentage") +
  ggtitle(yr)

ggplot(CData_yr, aes(x=source, y=area, fill=LC)) + 
  geom_bar(stat="identity", colour="white") +
  facet_grid(.~state) +
  xlab("Data Source")+ylab("Area km2") +
  ggtitle(yr)

```


Plot areas by year for a given state
```{r}
st <- "TO"

#filter (remove) missing LCs
CData_st <- CData %>%
  dplyr::filter(state == st) %>%
  dplyr::filter(area > 0)

ggplot(CData_st, aes(x=source, y=area, fill=LC)) + 
  geom_bar(stat="identity", colour="white", position = "fill") +
  scale_y_continuous(labels = percent_format()) +
  facet_grid(.~year) +
  xlab("Data Source")+ylab("Percentage") +
  ggtitle(st)

ggplot(CData_st, aes(x=source, y=area, fill=LC)) + 
  geom_bar(stat="identity", colour="white") +
  facet_grid(.~year) +
  xlab("Data Source")+ylab("Area km2") +
  ggtitle(st)

```

  ###Quantitative Differences
  CDataW <- CData %>%
    mutate(sourceLC = paste0(source, LC)) %>%
    dplyr::select(-source, -LC) %>%
    spread(key = c(sourceLC), value = area) %>%
    dplyr::select(state, starts_with("Map"), starts_with("MB"))
    
  kable(CDataW)
  
  #it may be that in a given classification some LCs may not exist. so to calculate summarise we need to do some dplyr programming
  
  #get the names of LCs that exist in this classification in a list
  d <- unlist(distinct(CData, LC), use.names=F)
  
  #iterate over the list of LC names
  #mustate lines with help from https://stackoverflow.com/a/26003971 and https://stackoverflow.com/a/44622499
  for(j in seq_along(d)){
    CDataW <- CDataW %>%
      mutate(!!paste0(d[j],"AbsDiffc") := .data[[paste0("MB",d[j])]] - .data[[paste0("Map",d[j])]]) %>%
      mutate(!!paste0(d[j],"PropDiffc") := round(.data[[paste0(d[j],"AbsDiffc")]] / .data[[paste0("MB",d[j])]],3)) 
  }

  ###Absolute Differences
  abs <- CDataW %>%
    dplyr::select(-starts_with("MB"), -starts_with("Map"), -ends_with("PropDiffc"))
  
  kable(abs)
  
  abs <- abs %>%
    gather(key = LC, value = area, -state)
  
  ggplot(abs, aes(x=state, y=area, fill=LC)) + 
    geom_bar(stat="identity", colour="white", position = "dodge") +
    xlab("Data Source")+ylab("Abs Diff km2")

  ###Proportional Diferences
  prop <- CDataW %>%
    dplyr::select(-starts_with("MB"), -starts_with("Map"), -ends_with("AbsDiffc")) 
  
  kable(prop)
  
  prop <- prop %>%
    gather(key = LC, value = area, -state)
  
  ggplot(prop, aes(x=state, y=area, fill=LC)) + 
    geom_bar(stat="identity", colour="white", position = "dodge") +
    xlab("Data Source")+ylab("Prop Diff")

  ##All States
  SDataW <- CData %>%
    mutate(sourceState = paste(source, state, sep=" ")) %>%
    dplyr::select(-source, -state) %>%
    spread(key = c(sourceState), value = area)
  
  #following with help from https://stackoverflow.com/a/42015507
  SDataW <- SDataW %>%
    mutate(MapTotal = rowSums(.[grep("Map", names(.))], na.rm =TRUE)) %>%
    mutate(MBTotal = rowSums(.[grep("MB", names(.))], na.rm =TRUE)) %>%
    dplyr::select(LC, MapTotal, MBTotal) %>%
    mutate(AbsDiffc = MBTotal - MapTotal) %>%
    mutate(PropDiffc = round(AbsDiffc / MBTotal,3))
  
  Stotals <- SDataW %>%
    dplyr::select(-AbsDiffc, -PropDiffc) %>%
    gather(key = source, value = area, -LC)
  
  ggplot(Stotals, aes(x=source, y=area, fill=LC)) + 
    geom_bar(stat="identity", colour="white", position = "fill") +
    scale_y_continuous(labels = percent_format()) +
    xlab("Data Source")+ylab("Percentage")
  
  ggplot(Stotals, aes(x=source, y=area, fill=LC)) + 
    geom_bar(stat="identity", colour="white") +
    xlab("Data Source")+ylab("Area km2")
  
  ggplot(SDataW , aes(x=LC, y=AbsDiffc)) + 
    geom_bar(stat="identity", colour="white", position = "dodge") +
    xlab("Data Source")+ylab("Abs Diff")
  
  ggplot(SDataW , aes(x=LC, y=PropDiffc)) + 
    geom_bar(stat="identity", colour="white", position = "dodge") +
    xlab("Data Source")+ylab("Prop Diff")
  
  MBTotalArea = sum(SDataW$MBTotal)
  MapTotalArea = sum(SDataW$MapTotal)
  ADiffc = MBTotalArea - MapTotalArea
  PDiffc = ADiffc / MBTotalArea
  
  SDataW <- rbind(SDataW, c('Total', MapTotalArea, MBTotalArea, ADiffc, round(PDiffc,3)))
  
  kable(SDataW)
  
  
  
  
  #last thing to do in the classifications loop is save the tables to a list
  CData_ls[[i]] <- CData
  CDataW_ls[[i]] <- CDataW
  SDataW_ls[[i]] <- SDataW
  Stotals_ls[[i]] <- Stotals



```

Now need code here to save all data tables to a list of data tables. 


##Clean up
```{r}

unlink("ASCII", recursive = T) #delete ASCII directory created above

```
