---
title: "MapBiomas Classification - Pasture Analysis"
author: "James D.A. Millington"
date: "July 2018"
output: 
  html_document: 
    keep_md: yes
    code_folding: hide
    df_print: paged
---

This script analyses pasture areas, meat production and dairy production (and intensities) for all states and years, and multiple classifications (see accompanying Excel file) for two data sources:

- _MB_ is the original 30m MapBiomas data
- _Map_ is our aggregated 5km map data


```{r results='hide', warning=F, message=F}

rm(list=ls())

packages <- c(
  "tidyverse",
  "raster",
  "readxl",       #for reading Excel sheets
  "scales",       #useful for ggplotting
  "knitr",
  "rasterVis",    #more useful raster plotting
  "cowplot"       #useful for ggplotting
  )     
  
#use lapply to suppress all wanings: https://stackoverflow.com/a/46685042
invisible(lapply(packages, function(xxx) suppressMessages(require(xxx, character.only = TRUE,quietly=TRUE,warn.conflicts = FALSE))))

```

```{r warning=F, message=F}

#raster to xyz  (with help from https://stackoverflow.com/a/19847419)
#sepcify input raster, whether nodata cells should be output, whether a unique cell ID should be added
#return is a matrix. note format is row (Y) then col (X)
extractXYZ <- function(raster, nodata = FALSE, addCellID = TRUE){
  
  vals <- raster::extract(raster, 1:ncell(raster))   #specify raster otherwise dplyr used
  xys <- rowColFromCell(raster,1:ncell(raster))
  combine <- cbind(xys,vals)
  
  if(addCellID){
    combine <- cbind(1:length(combine[,1]), combine)
  }
  
  if(!nodata){
    combine <- combine[!rowSums(!is.finite(combine)),]  #from https://stackoverflow.com/a/15773560
  }
  
  return(combine)
}


getLCs <- function(data)
{
  #calculates proportion of each LC in the muni (ignoring NAs, help from https://stackoverflow.com/a/44290753)
  data %>%
    group_by(muniID) %>%
    dplyr::summarise(LC1 = round(sum(map == 1, na.rm = T) / sum(!is.na(map)), 3),
                     LC2 = round(sum(map == 2, na.rm = T) / sum(!is.na(map)), 3),
                     LC3 = round(sum(map == 3, na.rm = T) / sum(!is.na(map)), 3),
                     LC4 = round(sum(map == 4, na.rm = T) / sum(!is.na(map)), 3),
                     LC5 = round(sum(map == 5, na.rm = T) / sum(!is.na(map)), 3),
                     NonNAs = sum(!is.na(map)),
                     NAs = sum(is.na(map))
    ) -> LCs

  return(LCs)
}


```


```{r results='hide', warning=F, messages=F}

unzip(zipfile="MapBiomas_23_ASCII_unclassified_allYears.zip")  # unzip all files 

#for 'suppressMessages' see https://stackoverflow.com/a/41229387
mb_data <- suppressMessages(read_csv("LandCover Data - MapBiomas - Collection 2.3 - 2018.01.04 Municipios.csv"))

unzip(zipfile="sim10_BRmunis_latlon_5km_2018-04-27.zip",files="sim10_BRmunis_latlon_5km_2018-04-27.asc",exdir="ASCII")  # unzip file 
munis.r <- raster("ASCII/sim10_BRmunis_latlon_5km_2018-04-27.asc")  #do this with zip file

#extract cell values to table format
munis.t <- extractXYZ(munis.r, addCellID = F)
munis.t <- as.data.frame(munis.t)
munis.t <- plyr::rename(munis.t, c("vals" = "muniID"))

```


```{r}
#Specify classifications and years to examine. Classifications should be the names of Sheets in the Classifications Excel file. Years should be between 2000 and 2015 

#classifications to loop through 
cls <- c("PastureA", "PastureB", "PastureC")

yrls <- seq(2000,2015,1)

```


```{r warning=F, message=F}

#lists to hold data tables 
CData_ls <- vector('list', length(cls))
CDataW_ls <- vector('list', length(cls))
SDataW_ls <- vector('list', length(cls))
Stotals_ls <- vector('list', length(cls))
SDataW_Adj_ls <- vector('list', length(cls))
mapStack_ls <- vector('list', length(cls))

names(CData_ls) <- cls
names(CDataW_ls) <- cls
names(SDataW_ls) <- cls
names(Stotals_ls) <- cls
names(SDataW_Adj_ls) <- cls
names(mapStack_ls) <- cls

#loop over classifications
for(i in seq_along(cls)){
  
  classification <- read_excel("MapBiomas_CRAFTY_classifications.xlsx", sheet = cls[i], range="B2:C21", col_names=F) 
  
  #reset mapStack for this Classification
  mapStack <- stack()

  #loop over years  
  for(j in seq_along(yrls)){

  #print(paste0("Classification: ",cls[i],", Year: ",yrls[j]))
  
  map <- raster(paste0("ASCII/brazillc_",yrls[j],"_5km_int.txt"))
  map <- reclassify(map, rcl=as.matrix(classification))
  
  
  #add categories for later plotting etc. (see https://stackoverflow.com/a/37214431)
  map <- ratify(map)     #tell R that the map raster is categorical 
  rat <- levels(map)[[1]]    #apply the levels (i.e. categories) 
 
  #not all classes may be present after classification, so conditionally construct labels
  labs <- c()
  if(1 %in% levels(map)[[1]]$ID) { labs <- c(labs, "Nature") }
  if(2 %in% levels(map)[[1]]$ID) { labs <- c(labs, "OtherAgri") }
  if(3 %in% levels(map)[[1]]$ID) { labs <- c(labs, "Agriculture") }
  if(4 %in% levels(map)[[1]]$ID) { labs <- c(labs, "Other") }
  if(5 %in% levels(map)[[1]]$ID) { labs <- c(labs, "Pasture") }
    
  rat$landcover <- labs  
  levels(map) <- rat 
  
  #add to mapStack for later plotting
  mapStack <- stack(map, mapStack)

  #extract cell values to table format
  map.t <- extractXYZ(map, addCellID = F)
  map.t <- as.data.frame(map.t)
  map.t <- plyr::rename(map.t, c("vals" = "map"))

  #so need to join 
  map_munis <- left_join(as.data.frame(munis.t), as.data.frame(map.t), by = c("row" = "row", "col" = "col"))

  #now summarise by muniID
  lcs_map_munis <- getLCs(map_munis)

  #convert cell counts to areas (km2) and add state id
  map_areas_munis <- lcs_map_munis %>%
    mutate(LC1area = round(LC1 * NonNAs) * 25) %>%
    mutate(LC2area = round(LC2 * NonNAs) * 25) %>%
    mutate(LC3area = round(LC3 * NonNAs) * 25) %>%
    mutate(LC4area = round(LC4 * NonNAs) * 25) %>%
    mutate(LC5area = round(LC5 * NonNAs) * 25) %>%
    mutate(state = substr(muniID, 1, 2))

  #drop original cell-count columns (work with area km2 from now on)
  map_areas_munis <- map_areas_munis %>% dplyr::select(-LC1, -LC2, -LC3, -LC4, -LC5, -NonNAs, -NAs)

  #summarise muni areas to state level
  map_areas <- map_areas_munis %>%
    group_by(state) %>%
    dplyr::summarise_at(vars(LC1area:LC5area),sum, na.rm=T) %>%  #use _at so state is not summarised
    mutate_if(is.character, as.integer)

  #gather to long format for union below
  map_areas <- map_areas %>%
    gather(key = LCa, value = area, -state)
  
  #recode LCs for union below
  map_areas <- map_areas %>%
    mutate(LC = if_else(LCa == "LC1area", 1, 
      if_else(LCa == "LC2area", 2,
      if_else(LCa == "LC3area", 3,
      if_else(LCa == "LC4area", 4,
      if_else(LCa == "LC5area", 5, 0)
      )))))

  #add source variable for plotting below (re-order to match map table for union below)
  map_areas <- map_areas %>%
    dplyr::select(-LCa) %>%
    mutate(source = "Map") %>%
    dplyr::select(state, LC, source, area)

  ###Summarise MapBiomas data to the states we are simulating

  #filter to get only the states we want
  mb_areas <- mb_data %>%
    filter(Estados == "TOCANTINS" | 
        Estados == "BAHIA" |
        Estados == "MINAS GERAIS" |
        Estados == "SÃO PAULO" |
        Estados == "PARANÁ" |
        Estados == "SANTA CATARINA" |
        Estados == "RIO GRANDE DO SUL" |
        Estados == "MATO GROSSO DO SUL" |
        Estados == "MATO GROSSO" |
        Estados == "GOIÁS")
  
  #add state column containing state ids
  mb_areas <- mb_areas %>%
    mutate(state = if_else(Estados == "TOCANTINS", 17, 
      if_else(Estados == "BAHIA", 29,
      if_else(Estados == "MINAS GERAIS", 31,
      if_else(Estados == "SÃO PAULO", 35,
      if_else(Estados == "PARANÁ", 41,
      if_else(Estados == "SANTA CATARINA", 42,
      if_else(Estados == "RIO GRANDE DO SUL", 43, 
      if_else(Estados == "MATO GROSSO DO SUL", 50, 
      if_else(Estados == "MATO GROSSO", 51,
      if_else(Estados == "GOIÁS", 52, 0
      ))))))))))
    )
  
      
  #select only columns we want    
  mb_areas <- mb_areas %>%
    dplyr::select(state, paste0(yrls[j]), `Classe Nivel 3`) %>% 
    dplyr::rename(area = paste0(yrls[j]))
  
  #because there is no consistency between land cover labels
  legenda = c("Forest Formations"=1, "Natural Forest Formations"=2, "Dense Forest"=3, "Savanna Formations"=4, "Mangroves"=5,"Forest Plantations"=9, "Non-Forest Natural Formations"=10, "Non Forest Wetlands"=11, "Grasslands"=12, "Other Non Forest Natural Formations"=13, "Farming"=14, "Pasture"=15, "Agriculture"=18, "Agriculture or Pasture"=21, "Non-Vegetated Areas"=22, "Dunes and Beaches"=23, "Urban Infrastructure"=24,"Other Non-Vegetated Area"=25,"Water Bodies"=26, "Non-Observed"=27)
  
  #recode to values (which match the map)
  mb_areas <- mb_areas %>%
    mutate(LC = recode(`Classe Nivel 3`, !!!legenda))
  
  #use the classification values from above to relassify land covers
  mb_areas$LC <- plyr::mapvalues(mb_areas$LC, from=as.numeric(classification$X__1), to=as.numeric(classification$X__2))
  
  #calculate total LC area by state
  mb_areas <- mb_areas %>%
    group_by(state, LC) %>%
    dplyr::summarise_at(vars(area),sum, na.rm=T) 
  
  #round to integer
  mb_areas <- mb_areas %>%
    mutate(area_km2 = round(area,0))
  
  #add source variable for plotting below (re-order to match map table for union below)
  mb_areas <- mb_areas %>%
    dplyr::select(-area) %>%
    mutate(source = "MB") %>%
    rename(area = area_km2) %>%
    dplyr::select(state, LC, source, area)
    

  CData_yr <- union_all(map_areas, mb_areas)

  #relabel states to characters
  CData_yr <- CData_yr %>%
    mutate(state = if_else(state == 17, "TO", 
      if_else(state == 29, "BA",
      if_else(state == 31, "MG",
      if_else(state == 35, "SP",
      if_else(state == 41, "PR",
      if_else(state == 42, "SC",
      if_else(state == 43, "RS", 
      if_else(state == 50, "MS",
      if_else(state == 51, "MT",
      if_else(state == 52, "GO", "NA"
      ))))))))))
    )
  
  #relabel LCs to characters
  CData_yr <- CData_yr %>%
    mutate(LC = if_else(LC == 1, "Nature", 
      if_else(LC == 2, "OtherAgri",
      if_else(LC == 3, "Agri",
      if_else(LC == 4, "Other",
      if_else(LC == 5, "Pasture", "NA"
      )))))
    )
  
  #add year column
  CData_yr <- CData_yr %>%
    mutate(year = yrls[j])
  
  #union CData for years here.
  #if first iteration of classification loop (re)create the tibble
  if(j == 1){
      CData <- CData_yr
  }
    
  #else join data to tibble (by creating another tibble, then join (ensure rows are not lost)
  else {
      CData <- union_all(CData, CData_yr)
  }
  }
  
  CData_ls[[i]] <- CData
  
  names(mapStack) <- yrls
  mapStack_ls[[i]] <- mapStack
}

```

#Maps
For 2000 for the three different classifications (for quick visual comparison)
```{r fig.width=12, warning=F, message=F}
for(i in seq_along(mapStack_ls)){
  
  clabs <- c()
  if(1 %in% levels(mapStack_ls[[i]])[[1]]$ID) { clabs <- c(clabs, 'forestgreen') }
  if(2 %in% levels(mapStack_ls[[i]])[[1]]$ID) { clabs <- c(clabs, 'wheat1') }
  if(3 %in% levels(mapStack_ls[[i]])[[1]]$ID) { clabs <- c(clabs, 'orange2') }
  if(4 %in% levels(mapStack_ls[[i]])[[1]]$ID) { clabs <- c(clabs, 'gray') }
  if(5 %in% levels(mapStack_ls[[i]])[[1]]$ID) { clabs <- c(clabs, 'darkolivegreen') }
  
  print(cls[i])

  print(rasterVis::levelplot(mapStack_ls[[i]]$X2000, pretty=T,att = 'landcover', col.regions=clabs, main=paste0(cls[[i]], " 2000")))
  
}

```


##Pasture area by state, year and classification
```{r warning=F, message=F}

#add classification lable to the CData tables
for(i in seq_along(cls)){
  
  CData_ls[[i]] <- CData_ls[[i]] %>%
    mutate(classification = cls[i])
}

#union the CData tables for the different classifications
for(i in seq_along(cls)){
  
  if(i == 1) { CDataU <- CData_ls[[i]] }
  else { CDataU <- dplyr::union(CDataU, CData_ls[[i]]) }
}

#filter to pasture only
CData_Pas <- CDataU %>%
  dplyr::filter(LC == "Pasture")
```


```{r fig.height=20, fig.width=12}

CData_Pas %>% 
  dplyr::filter(source == "Map") %>%
  ggplot(aes(x=classification, y=area, fill=classification)) + 
    geom_bar(stat="identity", colour="white", position = "dodge") +
    scale_y_continuous(name=expression(Area~km^{2}), labels = scales::comma) +
  #labs(x=expression(Data~Source),y=expression(Area~km^{2})) +
    facet_grid(year~state) +
    xlab("") +
    #ylab("Pasture Area km2") +
    theme(axis.text.x = element_blank()) +
  ggtitle("Map")

CData_Pas %>% 
  dplyr::filter(source == "MB") %>%
  ggplot(aes(x=classification, y=area, fill=classification)) + 
    geom_bar(stat="identity", colour="white", position = "dodge") +
    scale_y_continuous(name=expression(Area~km^{2}), labels = scales::comma) +
    facet_grid(year~state) +
    xlab("")+
    #ylab("Pasture Area km2") +
    theme(axis.text.x = element_blank()) +
  ggtitle("MB")

```


Table: Pasture areas, sorted on state and area
```{r rows.print=30}

arrange(CData_Pas, area, state, year)

```


```{r warning=F, message=F}
#Load Production Data
meat_prod_Astates <- read_excel("Cattle_meat_production_Kg_2000_2017_all_states.xlsx", sheet = "Plan1", skip = 1)  #data for all states Astates

#dairy data are by municiaplity for all states (Amunis)
dairy_prod_Amunis <- read_excel("dairy_Municipalities_Brazil.xlsx", sheet = "Tabela", skip = 1, na = c("", "-", "..."))
```

```{r echo=F, warning=F, message=F}
#Tidy production data
#tidy meat data
meat_prod_Astates %>%
  rename(state = NM_UF_SIGLA) %>%
  dplyr::select(-NM_UF, -CD_GCUF, -`2016`, -`2017`) %>%      #drop columns
  filter(!is.na(state)) %>%   #safer way to remove text line at bottom of state column
  mutate_at(vars("2000":"2015"), as.numeric) -> meat_prod_Astates

Fstate_vals <- c(17,	29,	31,	35,	41,	42,	43,	50,	51,	52)
Fstate_abbrev <- c("TO", "BA", "MG", "SP", "PR",  "SC", "RS", "MS", "MT", "GO")

#Filter meat focal states
meat_prod_Astates %>%
  filter(state %in% Fstate_abbrev)  -> meat_prod_Fstates

meat_prod_Fstates_long <- meat_prod_Fstates %>%
  gather(key = year, value = kg, -state) %>%
  mutate_at(vars(year), as.integer)


#dairy data need more work as are for municipalities
dairy_prod_Amunis <- dairy_prod_Amunis %>%
  rename(muniID = `IBGE CODE`) %>%
  filter(!is.na(muniID)) %>%   #safer way to remove text line in muniID 
  mutate(state = substr(muniID, 1, 2)) %>%     #extract the muniID
  mutate_at(vars("2000":"2015"), as.numeric) %>%  #convert values to numeric
  dplyr::select(-muniID, -Municipality)     #drop unwanted states 

dairy_prod_Amunis <- dairy_prod_Amunis %>%
  group_by(state) %>%  
  mutate_all(funs(. * 1.03 * 1000)) %>%     #convert from thousand litres to kgs
  summarise_all(sum, na.rm=T) -> dairy_Astates   #summarise munis to states

dairy_prod_Fstates<- dairy_prod_Amunis %>%
  filter(state %in% Fstate_vals) %>% 
  mutate(state=replace(state, 1:length(Fstate_vals), Fstate_abbrev))

dairy_prod_Fstates_long <- dairy_prod_Fstates %>%
  gather(key = year, value = kg, -state) %>%
  mutate_at(vars(year), as.integer)

```


```{r}

#join to pasture areas
meat_areas <- inner_join(CData_Pas, meat_prod_Fstates_long, by = c("year", "state"))
dairy_areas <- inner_join(CData_Pas, dairy_prod_Fstates_long, by = c("year", "state"))

#calculate intensities
meat_areas <- meat_areas %>%
  mutate(intensity = kg / area)

dairy_areas <- dairy_areas %>%
  mutate(intensity = kg / area)

```

#Meat
##Production by state, year and classification
Differentiated between original 30m MapBiomas reported data ( _MB_ ) and our aggregated 5km map data ( _map_ ) and

The next two plots have the same variables on the y-axis (Intensity) but the second plot is differently scaled (some bars are not plotted because their value is greater than the limit of the axis).
```{r fig.height=20, fig.width=12}

meat_areas %>% 
  filter(source == "Map") %>%
  ggplot(aes(x=classification, y=intensity, fill=classification)) + 
    geom_bar(stat="identity", colour="white", position = "dodge") +
    scale_y_continuous(name=expression(Intensity~kg~km^{2}),limits = c(0, 30000)) +
    facet_grid(year~state) +
    xlab("")+
    theme(axis.text.x = element_blank()) +
  ggtitle("Meat, Map")

meat_areas %>% 
  filter(source == "MB") %>%
  ggplot(aes(x=classification, y=intensity, fill=classification)) + 
    geom_bar(stat="identity", colour="white", position = "dodge") +
    scale_y_continuous(name=expression(Intensity~kg~km^{2}),limits = c(0, 30000)) +
    facet_grid(year~state) +
    xlab("") +
    theme(axis.text.x = element_blank()) +
  ggtitle("Meat, MB")

```

We see SC, SP and RS are all high relative to other states. We also see that intensities are generally greatest for _PatureA_ classification, lower for _PastureB_ and lowest for _PastureC_.  


##Summary stats over time
Now let's summarise the data over time and examine their means and medians (with variation - error bar is one SE)
```{r echo=F}
meat_areas %>%
  group_by(state, classification, source) %>%
  summarise(
    area_mn = round(mean(area),0),
    area_md = round(median(area),0),
    area_sd = round(sd(area),0),
    prod_mn = round(mean(kg),0),
    prod_md = round(median(kg),0),
    prod_sd = round(sd(kg),0),
    int_mn = round(mean(intensity),0),
    int_md = round(median(intensity),0),
    int_max = round(max(intensity),0),
    int_sd = round(sd(intensity),0),
    int_se = round(sd(intensity) / length(intensity), 3)
    ) -> meat_summary

```

```{r fig.width=8}
meat_summary %>%
  #filter(source == "MB") %>%
  ggplot(aes(x=classification, y=int_mn, fill=classification)) + 
  geom_errorbar(aes(ymin=int_mn-int_se, ymax=int_mn+int_se), width=.1) +
  geom_bar(stat="identity", colour="white", position = "dodge") +
  facet_grid(source~state) +
  scale_y_continuous(name=expression(Intensity~kg~km^{2}), labels = scales::comma) +
  xlab("")+
  theme(axis.text.x = element_blank()) +
  ggtitle("Meat, Mean")

meat_summary %>%
  #filter(source == "Map") %>%
  ggplot(aes(x=classification, y=int_md, fill=classification)) + 
  geom_errorbar(aes(ymin=int_md-int_se, ymax=int_md+int_se), width=.1) +
  geom_bar(stat="identity", colour="white", position = "dodge") +
  facet_grid(source~state) +
  scale_y_continuous(name=expression(Intensity~kg~km^{2}), labels = scales::comma) +
  xlab("")+
  theme(axis.text.x = element_blank()) +
  ggtitle("Meat, Median")

meat_summary %>%
  #filter(source == "Map") %>%
  ggplot(aes(x=classification, y=int_max, fill=classification)) + 
  #geom_errorbar(aes(ymin=int_md-int_se, ymax=int_md+int_se), width=.1) +
  geom_bar(stat="identity", colour="white", position = "dodge") +
  facet_grid(source~state) +
  scale_y_continuous(name=expression(Intensity~kg~km^{2}), labels = scales::comma) +
  xlab("")+
  theme(axis.text.x = element_blank()) +
  ggtitle("Meat, Maxima")
```

Again we see the same patterns; SC, SP and RS are greater than other states and _PastureA_ > _PastureB_ > _PastureC_.

If we plot again, but this time limiting the y-axis to 15,000 kg/km2 (max at 20,000)
```{r fig.width=8}
meat_summary %>%
  #filter(source == "MB") %>%
  ggplot(aes(x=classification, y=int_mn, fill=classification)) + 
  geom_errorbar(aes(ymin=int_mn-int_se, ymax=int_mn+int_se), width=.1) +
  geom_bar(stat="identity", colour="white", position = "dodge") +
  facet_grid(source~state) +
  scale_y_continuous(name=expression(Intensity~kg~km^{2}), labels = scales::comma,limits = c(0, 15000)) +
  xlab("")+
  theme(axis.text.x = element_blank()) +
  ggtitle("Meat, Mean")

meat_summary %>%
  #filter(source == "Map") %>%
  ggplot(aes(x=classification, y=int_md, fill=classification)) + 
  geom_errorbar(aes(ymin=int_md-int_se, ymax=int_md+int_se), width=.1) +
  geom_bar(stat="identity", colour="white", position = "dodge") +
  facet_grid(source~state) +
  scale_y_continuous(name=expression(Intensity~kg~km^{2}), labels = scales::comma,limits = c(0, 15000)) +
  xlab("")+
  theme(axis.text.x = element_blank()) +
  ggtitle("Meat, Median")

meat_summary %>%
  #filter(source == "Map") %>%
  ggplot(aes(x=classification, y=int_max, fill=classification)) + 
  #geom_errorbar(aes(ymin=int_md-int_se, ymax=int_md+int_se), width=.1) +
  geom_bar(stat="identity", colour="white", position = "dodge") +
  facet_grid(source~state) +
  scale_y_continuous(name=expression(Intensity~kg~km^{2}), labels = scales::comma,limits = c(0, 20000)) +
  xlab("")+
  theme(axis.text.x = element_blank()) +
  ggtitle("Meat, Maxima")
```

We can now more clearly compare classifications. In some states (PR, SP) _PastureA_ and _PastureB_ are very similar values, but in many others _PastureB_ is lower; _PastureC_ is noticably smaller than the other classifications. We also see that Intensity is greater for  _MB_ data compared to _Map_. (NB: _PastureB_ for SC is very similar for MB data)

**Seems feasible to use _PastureB_ classification?**

If so, the 'perfect' meat intensity would be ~9,200 kg km2 if using mean/median or ~11,000 if using max (see raw tables of summary data below)

```{r}

meat_summary %>%
  filter(source == "Map" & classification == "PastureB") %>%
  dplyr::select(-starts_with("prod"), -int_sd, -int_se) %>%
  filter(state != "SC") 

```


Or, mean of maxima (without SC) is ~4,500

```{r}

meat_summary %>%
  filter(source == "Map" & classification == "PastureB") %>%
  dplyr::select(-starts_with("prod"), -int_sd, -int_se) %>%
  filter(state != "SC") %>%
  group_by(classification) %>%
  summarise_at(vars(int_max), mean)

```

In discussion with Ramon (email 06/07/18): 

"the pasture B seems very reasonable to me. More realistic than the other situations.
To make sure our assumptions based on the results, I was checking some data from the livestock sector and I found [this report](http://csr.ufmg.br/pecuaria/portfolio-item/cenarios-para-o-brasil/). All in Portuguese, but the tables and figures is not hard to understand.

The point is: based on future scenarios to the livestock sector in Brazil, they expect a productivity of 87.3kg/hectare in 2030 if the Brazilian sector modernize. It means that the potential is that the same as exactly the "perfect" meat intensity that you found based on the mapabiomas data (that is 9,100kg/km2).

*See the table row "Produção de @/hectare". @ is equivalent to 15kg."

87.3 kg/ha == 8,730 kg/sq km

This is comforting as it is the same order of magnitude as we found above. **Given we want some indication of 'perfect' we will use the higher value of 11,000 kg/sq km (the maximum of all states bar SC in PastureB land cover re-classification**


In turn, 11,000 kg/sq km == **0.275 gg/25sq km**

Full meat summary data table for reference:

```{r }

kable(arrange(meat_summary, desc(int_mn), state, classification), caption="Meat data, sorted on mean production values descending")

```

#Dairy
##Production by state, year and classification
Differentiated between original 30m MapBiomas data ( _MB_ ) and our aggregated 5km map data ( _map_ ).

```{r}

dairy_areas %>%
  group_by(state, classification, source) %>%
  summarise(
    area_mn = round(mean(area),0),
    area_md = round(median(area),0),
    area_sd = round(sd(area),0),
    prod_mn = round(mean(kg),0),
    prod_md = round(median(kg),0),
    prod_sd = round(sd(kg),0),
    int_mn = round(mean(intensity),0),
    int_md = round(median(intensity),0),
    int_max = round(max(intensity),0),
    int_sd = round(sd(intensity),0),
    int_se = round(sd(intensity) / length(intensity), 3)
    ) -> dairy_summary

```


```{r fig.height=20, fig.width=12}

dairy_areas %>% 
  filter(source == "Map") %>%
  ggplot(aes(x=classification, y=intensity, fill=classification)) + 
    geom_bar(stat="identity", colour="white", position = "dodge") +
    scale_y_continuous(name=expression(Intensity~kg~km^{2}), labels = scales::comma) +
    facet_grid(year~state) +
    xlab("") +
    theme(axis.text.x = element_blank()) +
  ggtitle("Dairy, Map")

dairy_areas %>% 
  filter(source == "MB") %>%
  ggplot(aes(x=classification, y=intensity, fill=classification)) + 
    geom_bar(stat="identity", colour="white", position = "dodge") +
    scale_y_continuous(name=expression(Intensity~kg~km^{2}), labels = scales::comma) +
    facet_grid(year~state) +
    xlab("") +
    theme(axis.text.x = element_blank()) +
  ggtitle("Dairy, MB")

```

The magnitude of Intensity for dairy in SC (and to some extent RS) is even more stark than for meat production, especially for our _Map_ data.   

Let's not re-scale the y-axis here, and instead go direct to the summary values. 

##Summary stats over time
```{r fig.width=8}
dairy_summary %>%
  #filter(source == "MB") %>%
  ggplot(aes(x=classification, y=int_mn, fill=classification)) + 
  geom_errorbar(aes(ymin=int_mn-int_se, ymax=int_mn+int_se), width=.1) +
  geom_bar(stat="identity", colour="white", position = "dodge") +
  facet_grid(source~state) +
  scale_y_continuous(name=expression(Intensity~kg~km^{2}), labels = scales::comma) +
  xlab("")+
  theme(axis.text.x = element_blank()) +
  ggtitle("Dairy, Mean")

dairy_summary %>%
  #filter(source == "Map") %>%
  ggplot(aes(x=classification, y=int_md, fill=classification)) + 
  geom_errorbar(aes(ymin=int_md-int_se, ymax=int_md+int_se), width=.1) +
  geom_bar(stat="identity", colour="white", position = "dodge") +
  facet_grid(source~state) +
  scale_y_continuous(name=expression(Intensity~kg~km^{2}), labels = scales::comma) +
  xlab("")+
  theme(axis.text.x = element_blank()) +
  ggtitle("Dairy, Median")

dairy_summary %>%
  #filter(source == "Map") %>%
  ggplot(aes(x=classification, y=int_max, fill=classification)) + 
  #geom_errorbar(aes(ymin=int_md-int_se, ymax=int_md+int_se), width=.1) +
  geom_bar(stat="identity", colour="white", position = "dodge") +
  facet_grid(source~state) +
  scale_y_continuous(name=expression(Intensity~kg~km^{2}), labels = scales::comma) +
  xlab("")+
  theme(axis.text.x = element_blank()) +
  ggtitle("Dairy, Maxima")

```

Again, we see the very high intensities for SC, RS and PR. 

Plot the summary stats again, but this time limit the y-axis to 250,000 kg km2
```{r fig.width=8}
dairy_summary %>%
  #filter(source == "MB") %>%
  ggplot(aes(x=classification, y=int_mn, fill=classification)) + 
  geom_errorbar(aes(ymin=int_mn-int_se, ymax=int_mn+int_se), width=.1) +
  geom_bar(stat="identity", colour="white", position = "dodge") +
  facet_grid(source~state) +
  scale_y_continuous(name=expression(Intensity~kg~km^{2}), labels = scales::comma, limits = c(0, 250000)) +
  xlab("")+
  theme(axis.text.x = element_blank()) +
  ggtitle("Dairy, Mean")

dairy_summary %>%
  #filter(source == "Map") %>%
  ggplot(aes(x=classification, y=int_md, fill=classification)) + 
  geom_errorbar(aes(ymin=int_md-int_se, ymax=int_md+int_se), width=.1) +
  geom_bar(stat="identity", colour="white", position = "dodge") +
  facet_grid(source~state) +
  scale_y_continuous(name=expression(Intensity~kg~km^{2}), labels = scales::comma, limits = c(0, 250000)) +
  xlab("")+
  theme(axis.text.x = element_blank()) +
  ggtitle("Dairy, Median")

dairy_summary %>%
  #filter(source == "Map") %>%
  ggplot(aes(x=classification, y=int_max, fill=classification)) + 
  #geom_errorbar(aes(ymin=int_md-int_se, ymax=int_md+int_se), width=.1) +
  geom_bar(stat="identity", colour="white", position = "dodge") +
  facet_grid(source~state) +
  scale_y_continuous(name=expression(Intensity~kg~km^{2}), labels = scales::comma, limits = c(0, 250000)) +
  xlab("")+
  theme(axis.text.x = element_blank()) +
  ggtitle("Dairy, Maxima")
```

_PastureA_ for SC is not plotted at its value is so large. 

**I think we maybe need to ignore SC as it has such high intensity?**  (High intensity is due to very low area - see plots at top, and raw areas table at bottom, of this file)

**And we should use the** _PastureB_ **classification?**

If we ignore SC and use _PastureB_ the next largest _PastureB_ dairy intensity value (for Map) is PR (~67,000 for mean/median, ~98,000 for max; see Table below)

```{r}

dairy_summary %>%
  filter(source == "Map" & classification == "PastureB") %>%
  dplyr::select(-starts_with("prod"), -int_sd, -int_se) %>%
  filter(state != "SC") 

```


Or, mean of the maxima (excluding SC) is ~25,000

Table: Summary Dairy data _PastureB_ classification for _Map_ 
```{r}

dairy_summary %>%
  filter(source == "Map" & classification == "PastureB") %>%
  dplyr::select(-starts_with("prod"), -int_sd, -int_se) %>%
  filter(state != "SC") %>%
  group_by(classification) %>%
  summarise_at(vars(int_max), mean)

```


In discussion with Ramon (email 10/07/18):

"about the data to support the model results for dairy production, probably will not be "simple" as we found to the meat. For dairy production there are many variables and land is just one. Technology and feed supply are just two to mention. We know that the same variables are important for meat but in dairy production is higher significant. Other issue is that for dairy production the productivity is usually measuredTherefore,  by other ways, such as liters/number of animals. However, I am sure taht we can find data to better help us to understand the model's result.

Therefore, based on [a study from Embrapa](https://www.infoteca.cnptia.embrapa.br/bitstream/doc/890118/1/Areasdeconcentracao.pdf), in 2006 the higher productivity was 89 thousand liters of milk per square.kilometer.

Converting to liters to kg:

89,000 * 1.03 = 91,670

As above for meat, this is comforting as it shows the correct order of magnitde. Again, **given we want some indication of 'perfect' we will use the higher value of 100,000 kg/sq km (the maximum of all states bar SC in PastureB land cover re-classification**


In turn, 100,000 kg/sq km == **2.5 gg/25sq km**

Full table for reference below.

Table: Dairy data, sorted on mean values descending
```{r rows.print=15}

arrange(dairy_summary, desc(int_mn), state, classification)

```



```{r}

unlink("ASCII", recursive = T) #delete ASCII directory created above

```
