---
title: "MapBiomasClassification"
output: 
  html_document: 
    keep_md: yes
---

##Load Libraries
```{r warning=F, message=F}

rm(list=ls())
library(tidyverse)
library(raster)
library(readxl)
library(scales)

```

##Functions
```{r}

#raster to xyz  (with help from https://stackoverflow.com/a/19847419)
#sepcify input raster, whether nodata cells should be output, whether a unique cell ID should be added
#return is a matrix. note format is row (Y) then col (X)
extractXYZ <- function(raster, nodata = FALSE, addCellID = TRUE){
  
  vals <- raster::extract(raster, 1:ncell(raster))   #specify raster otherwise dplyr used
  xys <- rowColFromCell(raster,1:ncell(raster))
  combine <- cbind(xys,vals)
  
  if(addCellID){
    combine <- cbind(1:length(combine[,1]), combine)
  }
  
  if(!nodata){
    combine <- combine[!rowSums(!is.finite(combine)),]  #from https://stackoverflow.com/a/15773560
  }
  
  return(combine)
}


getLCs <- function(data)
{
  #calculates proportion of each LC in the muni (ignoring NAs, help from https://stackoverflow.com/a/44290753)
  data %>%
    group_by(muniID) %>%
    dplyr::summarise(LC1 = round(sum(map == 1, na.rm = T) / sum(!is.na(map)), 3),
                     LC2 = round(sum(map == 2, na.rm = T) / sum(!is.na(map)), 3),
                     LC3 = round(sum(map == 3, na.rm = T) / sum(!is.na(map)), 3),
                     LC4 = round(sum(map == 4, na.rm = T) / sum(!is.na(map)), 3),
                     LC5 = round(sum(map == 5, na.rm = T) / sum(!is.na(map)), 3),
                     NonNAs = sum(!is.na(map)),
                     NAs = sum(is.na(map))
    ) -> LCs

  return(LCs)
}


```

##Load Data
Maps from Zip
```{r}

unzip(zipfile="MapBiomas_23_ASCII_unclassified_allYears.zip",files="ASCII/brazillc_2000_5km_int.txt")  # unzip file 
map <- raster("ASCII/brazillc_2000_5km_int.txt") 
plot(map) 

```


Classification from Excel
```{r warning=F, message=F}

classification <- read_excel("MapBiomas_CRAFTY_classifications.xlsx", sheet = "PastureB", range="B2:C21", col_names=F)  

```

MapBiomas land areas from csv (areas are km2)
```{r warning=F, message=F}

mb_data <- read_csv("LandCover Data - MapBiomas - Collection 2.3 - 2018.01.04 Municipios.csv")  


```


##Classify
```{r}

map <- reclassify(map, rcl=as.matrix(classification))
plot(map, col=c("green", "pink", "yellow", "grey", "brown"))

```


##Compare map areas to MapBiomas area 
Let's compare areas at the state level for the 10 states being simulated. These are:
IBGE CODE	STATE
17	Tocantins
29	Bahia
31	Minas Gerais
35	São Paulo
41	Paraná
42	Santa Catarina
43	Rio Grande do Sul
50	Mato Grosso do Sul
51	Mato Grosso
52	Goiás

Create a summary table of classified map data for municipalities in states we are simulating.
```{r}

unzip(zipfile="sim10_BRmunis_latlon_5km_2018-04-27.zip",files="sim10_BRmunis_latlon_5km_2018-04-27.asc",exdir="ASCII")  # unzip file 
munis.r <- raster("ASCII/sim10_BRmunis_latlon_5km_2018-04-27.asc")  #do this with zip file

#extract cell values to table format
munis.t <- extractXYZ(munis.r, addCellID = F)
map.t <- extractXYZ(map, addCellID = F)

munis.t <- as.data.frame(munis.t)
munis.t <- plyr::rename(munis.t, c("vals" = "muniID"))


map.t <- as.data.frame(map.t)
map.t <- plyr::rename(map.t, c("vals" = "map"))

#so need to join 
map_munis <- left_join(as.data.frame(munis.t), as.data.frame(map.t), by = c("row" = "row", "col" = "col"))

#now summarise by muniID
lcs_map_munis <- getLCs(map_munis)

#convert cell counts to areas (km2) and add state id
map_areas_munis <- lcs_map_munis %>%
  mutate(LC1area = round(LC1 * NonNAs) * 25) %>%
  mutate(LC2area = round(LC2 * NonNAs) * 25) %>%
  mutate(LC3area = round(LC3 * NonNAs) * 25) %>%
  mutate(LC4area = round(LC4 * NonNAs) * 25) %>%
  mutate(LC5area = round(LC5 * NonNAs) * 25) %>%
  mutate(state = substr(muniID, 1, 2))

#drop original cell-count columns (work with area km2 from now on)
map_areas_munis <- map_areas_munis %>% dplyr::select(-LC1, -LC2, -LC3, -LC4, -LC5, -NonNAs, -NAs)

#summarise muni areas to state level
map_areas <- map_areas_munis %>%
  group_by(state) %>%
  dplyr::summarise_at(vars(LC1area:LC5area),sum, na.rm=T) %>%  #use _at so state is not summarised
  mutate_if(is.character, as.integer)

#gather to long format for union below
map_areas <- map_areas %>%
  gather(key = LCa, value = area, -state)

#recode LCs for union below
map_areas <- map_areas %>%
  mutate(LC = if_else(LCa == "LC1area", 1, 
    if_else(LCa == "LC2area", 2,
    if_else(LCa == "LC3area", 3,
    if_else(LCa == "LC4area", 4,
    if_else(LCa == "LC5area", 5, 0)
    )))))

#add source variable for plotting below (re-order to match map table for union below)
map_areas <- map_areas %>%
  dplyr::select(-LCa) %>%
  mutate(source = "Map") %>%
  dplyr::select(state, LC, source, area)
      
```


Summarise MapBiomas data to the states we are simulating
```{r}

#filter to get only the states we want
mb_areas <- mb_data %>%
  filter(Estados == "TOCANTINS" | 
      Estados == "BAHIA" |
      Estados == "MINAS GERAIS" |
      Estados == "SÃO PAULO" |
      Estados == "PARANÁ" |
      Estados == "SANTA CATARINA" |
      Estados == "RIO GRANDE DO SUL" |
      Estados == "MATO GROSSO DO SUL" |
      Estados == "MATO GROSSO" |
      Estados == "GOIÁS")

#add state column containing state ids
mb_areas <- mb_areas %>%
  mutate(state = if_else(Estados == "TOCANTINS", 17, 
    if_else(Estados == "BAHIA", 29,
    if_else(Estados == "MINAS GERAIS", 31,
    if_else(Estados == "SÃO PAULO", 35,
    if_else(Estados == "PARANÁ", 41,
    if_else(Estados == "SANTA CATARINA", 42,
    if_else(Estados == "RIO GRANDE DO SUL", 43, 
    if_else(Estados == "MATO GROSSO DO SUL", 50, 
    if_else(Estados == "MATO GROSSO", 51,
    if_else(Estados == "GOIÁS", 52, 0
    ))))))))))
  )

    
#select only columns we want    
mb_areas <- mb_areas %>%
  dplyr::select(state, `2000`, `Classe Nivel 3`) %>% 
  dplyr::rename(area = `2000`)

#because there is no consistency between land cover labels
legenda = c("Forest Formations"=1, "Natural Forest Formations"=2, "Dense Forest"=3, "Savanna Formations"=4, "Mangroves"=5,"Forest Plantations"=9, "Non-Forest Natural Formations"=10, "Non Forest Wetlands"=11, "Grasslands"=12, "Other Non Forest Natural Formations"=13, "Farming"=14, "Pasture"=15, "Agriculture"=18, "Agriculture or Pasture"=21, "Non-Vegetated Areas"=22, "Dunes and Beaches"=23, "Urban Infrastructure"=24,"Other Non-Vegetated Area"=25,"Water Bodies"=26, "Non-Observed"=27)

#recode to values (which match the map)
mb_areas <- mb_areas %>%
  mutate(LC = recode(`Classe Nivel 3`, !!!legenda))

#use the classification values from above to relassify land covers
mb_areas$LC <- plyr::mapvalues(mb_areas$LC, from=as.numeric(classification$X__1), to=as.numeric(classification$X__2))

#calculate total LC area by state
mb_areas <- mb_areas %>%
  group_by(state, LC) %>%
  dplyr::summarise_at(vars(area),sum, na.rm=T) 

#round to integer
mb_areas <- mb_areas %>%
  mutate(area_km2 = round(area,0))

#add source variable for plotting below (re-order to match map table for union below)
mb_areas <- mb_areas %>%
  dplyr::select(-area) %>%
  mutate(source = "MB") %>%
  rename(area = area_km2) %>%
  dplyr::select(state, LC, source, area)
  
#mb_areas <- mb_areas %>%
#  spread(LC, area_km2) %>%
#  rename(MB1 = `1`, MB2 = `2`, MB3 = `3`, MB4 = `4`, MB5 = `5`)

```


Finally, join the MapBiomas data to the map data
```{r}

CData <- union_all(map_areas, mb_areas)


```

##Plotting
```{r}



ggplot(CData, aes(x=source, y=area, fill=factor(LC))) + 
  geom_bar(stat="identity", colour="white", position = "fill") +
  scale_y_continuous(labels = percent_format()) +
  facet_grid(.~state) +
  xlab("Beverage")+ylab("Mean Percentage")


ggplot(means.long,aes(x=variable,y=value,fill=factor(gender)))+
  geom_bar(stat="identity",position="dodge")+
  scale_fill_discrete(name="Gender",
                      breaks=c(1, 2),
                      labels=c("Male", "Female"))+
  xlab("Beverage")+ylab("Mean Percentage")



```



##Clean up
```{r}

#unlink("ASCII", recursive = T) #delete ASCII directory created above

```
